---
title: "Math-331 Project 2"
output: html_document
date: "2022-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Understanding the factors on which the pricing of a car depends in the market

### Objectives- 
1. Which variables are significant in predicting the price of a car
2. How well those variables describe the price of a car

### Importing the packages required

```{r}
library(ggplot2) 
library(tidyverse)
library(dplyr)
library(stringr)
library(class)
library(lubridate)
library(splines) #Regression spline functions and classes.
library(mgcv) #Generalized additive models.
library(randomForest) #Model the training and test sets
library(rpart) #Used for building classification and regression trees
library(rpart.plot)
```

### Importing the CarPrice dataset which is a large dataset of different types of cars across the American market.
```{r}
carPrice<- read.csv("/Users/malaika/Desktop/MATH-331/CarPrice.csv")
```

### Viewing the dataset
```{r}
#carPrice
```

### Cleaning the dataset

Removing the column car_ID as it does not contribute to our analysis
```{r}
carPrice<-subset(carPrice[,-1])
#carPrice
```

Splitting the CarName column into two columns: Car Company and Car Model 
```{r}
carPrice<- (carPrice %>% 
  separate(CarName, c('CarCompany', 'CarModel'),' ')%>% 
    drop_na())
```

### There are 21 columns (variables) in this data set of different data types: integer, character & numeric

After analyzing the table I found some spelling errors in the CarCompany column that need to be corrected

Using unique() to find unique values within the column
```{r}
unique(carPrice$CarCompany)
```

Correcting the spelling erros within the CarCompany column
```{r}
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "maxda", "mazda")
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "Nissan", "nissan")
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "porcshce", "porsche")
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "toyouta", "toyota")
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "vokswagen", "volkswagen")
carPrice$CarCompany <- str_replace(carPrice$CarCompany, "vw", "volkswagen")

#carPrice
```

Changing the data types of some of the variables to be able to analyze and manipulate them correctly

```{r}
carPrice$symboling <- as.factor(carPrice$symboling)
carPrice$CarCompany<-as.factor(carPrice$CarCompany)
carPrice$fueltype<-as.factor(carPrice$fueltype)
carPrice$aspiration<-as.factor(carPrice$aspiration)
carPrice$doornumber<-as.factor(carPrice$doornumber)
carPrice$carbody<-as.factor(carPrice$carbody)
carPrice$drivewheel<-as.factor(carPrice$drivewheel)
carPrice$enginelocation<-as.factor(carPrice$enginelocation)
carPrice$enginetype<-as.factor(carPrice$enginetype)
carPrice$cylindernumber<-as.factor(carPrice$cylindernumber)
carPrice$fuelsystem<-as.factor(carPrice$fuelsystem)
```

```{r}
#carPrice
```
I wasn't sure what the symboling variable meant so I made a plot to see how the symboling and price were related:
```{r}
ggplot(data=carPrice, aes(x=symboling,y=price,fill="red")) + 
  ylab("price")+
  geom_boxplot() + 
  ggtitle("Symboling in relation to price")+
  labs(caption = "Figure 1")
```
* From the above plots the symboling with 0 and 1 values have a high number of rows which means they are sold the most.

* The cars with -1 symboling have the highest price  

* Cars with symboling 3 have the price range similar to -2. There is a dip in price at symboling 1.

### Training the model
Since we're creating a predictive model we need to split the data into a train and test set. This allows our model to categorize and generalize data it has not worked on before instead of using a specific function that will end up being an overfit (over accurate). By first using the testing set, the model can identify a pattern and then apply it on the testing set to confirm if the model has accurately recognized a generalized pattern. 

Since my response variable is numeric I will be using regression and not classification models.

### Train/test/split the data using a 60/40 random split:
```{r}
n <- round( 0.6 * nrow(carPrice)) #no of rows for the sample set=60% 
  in_train <- sample(1:nrow(carPrice),n) #sample not the rows directly but draw a sample of the row numbers
#select the rows and columns
  train <- carPrice[in_train,]#rows
  test <- carPrice[-in_train,]#not the rows
  rm(carPrice,in_train)#remove extra variables added and remove carPrice since we've split our dataset 
  #train2<-train %>% select(-name,-category)
  
#write sets to files
write.csv(train,"train.csv",row.names =FALSE)
write.csv(test,"test.csv",row.names=FALSE)
```

### Decision Tree 
A decision tree is a supervised ML algorithm
It follows a set of if-else conditions 
We can visualize the data and classify it according to the conditions
```{r}
tree.model <- rpart(price ~ . -CarModel, data=train)
rpart.plot(tree.model)
```

I summarized the train set, and use that information to determine which factor variables are important in predicting the response.
```{r}
summary(train)
```

Creating a linear model with price as the response variable and all others as predictors. 
```{r}
my.mod<-lm(price~ .-1 -CarModel, data=train)
summary(my.mod)
```

```{r}
my.mod<-lm(price~ .-1 -CarModel, data=test)
summary(my.mod)
```

### Random Forest Model

Training the CarPrice dataset using the random forest model.This model builds decision trees on different samples and takes their average in case of regression and provides the highest accuracy.
```{r}
rf.model <- randomForest(price ~ .-CarModel, data=train,
                         ntree=1500,
                         mtry=5)
                        
predictions <- predict(rf.model)
```

Plotting the random forest model
```{r}
plot(rf.model)
```

```{r}
sum((predictions-train$price)^2)/length(predictions)
```

Assess on the test set: Assess the model performance on the test set by retraining the model using the parameter values and the entire train set. 
```{r}
predictions <- predict(rf.model, newdata=test)
sum((predictions-test$price)^2)/length(predictions)
```

```{r}
cor(predictions, test$price)^2
```


Finding the most important variables in the model:
```{r}
varImpPlot(rf.model)
```

